### 2013-11-12 paper/analysis.tex
Fix figure paths

\subsection{Statistical and sensitivity analysis}
\label{sec:forward}
We now exploit the PC surrogate to study the statistics of the 
water surface elevation, and to quantify its sensitivity to the
uncertain input parameters and thus to anticipate their impact on
the inverse problem.  In doing so, it is emphasized that 
no additional Geoclaw simulations were needed to obtain
the information presented below, rather it is obtained either directly 
from the coefficients of the PC surrogate or by sampling the corresponding
representations for different values of $\xxi$.

Figure~\ref{fig:ave} shows the evolution of
PC mean water surface elevation and its two standard deviations
bounds at the four gauge locations as indicated in each panel.  
The standard deviation is insignificant during the two hours
of the simulations and therefore was not included in the plots.
This standard deviation however increases as the tsunami increase
and its effect approaches the gauge locations. 
To quantity the contribution of each
uncertain parameter to the variance in water surface elevation, we calculate the total sensitivity index 
using the PC coefficients ~\citep{Alexanderian2012,Sudret,Crestaux}. The total sensitivity index
of  each of the uncertain parameters is shown in
Figure~\ref{fig:sens} for the four gauges. The Manning's roughness coefficient
at the shore $N_2$ is clearly dominant and contributes
most to the  water surface elevation variance compared to the other two 
Manning's roughness coefficients;
this is true throughout the simulation time.  The Manning's roughness coefficient
in the ocean $N_{3}$ at gauge number 21419 exhibits small sensitivity index 
the during the second hour of simulation and Manning's roughness coefficient
in the land $N_1$ appears to be an insignificant contributor
to the variance.
\begin{figure}[h]
\begin{tabular}{clc}

\includegraphics[width=0.475\textwidth]{../figures/musigma1.pdf} &
\includegraphics[width=0.475\textwidth]{../figures/musigma2.pdf} \\
\includegraphics[width=0.475\textwidth]{../figures/musigma3.pdf} &
\includegraphics[width=0.475\textwidth]{../figures/musigma4.pdf}
\end{tabular}
\caption{Evolution of PC mean water surface elevation at different gauge locations.}
\label{fig:ave}
\end{figure}
\begin{figure}[h]
\begin{tabular}{clc}
\includegraphics[width=0.475\textwidth]{../figures/sens1.pdf} &
\includegraphics[width=0.475\textwidth]{../figures/sens2.pdf} \\
\includegraphics[width=0.475\textwidth]{../figures/sens3.pdf} &
\includegraphics[width=0.475\textwidth]{../figures/sens4.pdf}
\end{tabular}
\caption{Total sensitivity index of different input parameters.}
\label{fig:sens}
\end{figure}

%The impact of $\alpha$ and $V_{max}$ for fixed $m=0$ is illustrated in Figure \ref{fig:response}
%for the area-averaged SST within the $42~km\times42~km$ analysis region; the
%different panels illustrate the time evolution of the SST response
%surfaces. Their most striking features are the relatively flat
%horizontal contours on Sep~17 and Sep~18 indicating that SST depends
%only mildly on $V_{\max}$ even during peak winds;
% unsurprisingly these contours turn completely flat on Sep
%19 (and afterwards) when the winds dip below 10~m/s. The multiplicative
%factor $\alpha$, however, exerts a strong influence on SST even
%under mild wind conditions.  This influence is relatively
%weak for the low $\alpha$ range and increases substantially for the
%higher $\alpha$ as evidenced by the packed contours.  The temperature
%response surfaces at 50-m and 200-m depths (not shown) exhibit
%roughly the same structure as the surface, with milder
%dependence on $V_{\max}$ even during the peak winds of Sep~18.

The same statistical analysis can be performed for the
entire domain in 2D. Figure~\ref{fig:mean2d}(top row) shows
the PC mean water surface elevation for the considered computational
domain at three different times as indicated in the title of each panel.
The standard deviation is also shown in Figure~\ref{fig:mean2d}(bottom row).
\begin{figure}[h]
        \begin{tabular}{ccc}
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/mean2d1.pdf} &
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/mean2d3.pdf} &
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/mean2d4.pdf} \\
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/sigma2d1.pdf} &
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/sigma2d3.pdf} &
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/sigma2d4.pdf}
\end{tabular}
\caption{PC mean (top row) and standard deviation (bottom row) of the water surface elevation at different times.}
\label{fig:mean2d}
\end{figure}

\begin{figure}[h]
\begin{tabular}{clc}
 \hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/T12d1.pdf} &
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/T12d3.pdf} &
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/T12d4.pdf} \\
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/T22d1.pdf} &
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/T22d3.pdf} &
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/T22d4.pdf} \\
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/T32d1.pdf} &
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/T32d3.pdf} &
\hspace*{-65pt}
\includegraphics[width=0.45\textwidth]{../figures/T32d4.pdf}
\end{tabular}
\caption{Total sensitivity index for $N_1$ (top row) $N_2$ (center row) and $N_3$ (bottom row)
 at different times as indicated.}
\end{figure}

>>>>

\subsection{Forward Propagation of Uncertainty}
\label{sec:forward}
We now use the surrogate created effectively using PC to propagate prior 
uncertainty through the forward model. We mainly exploit the surrogate 
to study the statistics of the water surface elevation, and to conduct 
a global sensitivity analysis of the impact of the uncertain input parameters.
We note that the statistical and sensitivity analysis presented below 
is performed using the PC coefficients solely; these were calculated
from an ensemble of 125 Geoclaw simulations.
\subsubsection{Statistical analysis}
The mean of a QoI and its standard deviation is computed
from the PC coefficients as shown in Equation~\ref{eq:mean} and Equation~\ref{eq:sigma}. 
In Figure~\ref{fig:ave} we plot the evolution of
the calculated PC mean water surface elevation along with two standard deviations
bounds at the four gauges as indicated in each panel.  
Note that we only show the evolution after $t=6000~s$ where the uncertainty is significant
reflected in a significant standard deviation as seen earlier from the realizations in Figure~\ref{fig:rlzs}.
We observed that the standard deviation ......

\comment{The same statistical analysis can be performed for the
entire domain in 2D. Figure~\ref{fig:mean2d}(top row) shows
the PC mean water surface elevation for the considered computational
domain at three different times as indicated in the title of each panel.
The standard deviation is also shown in Figure~\ref{fig:mean2d}(bottom row).}

\subsubsection{Sensitivity analysis}
A global sensitivity analysis is next conducted to quantity the contribution of each
uncertain parameter to the variance in water surface elevation. To this end, we calculate 
the total sensitivity index using the PC coefficients as shown in Equation~\ref{eq:sens}~\citep{Alexanderian2012,Sudret,Crestaux}. The evolution of the total sensitivity index
of each of the uncertain parameters is shown in Figure~\ref{fig:sens} for the four gauges. 
The Manning's roughness coefficient at the shore $N_2$ is clearly dominant and contributes
the most to the variance in the water surface elevation compared to the other two 
Manning's roughness coefficients $N_1$ and $N_3$; this is true for almost the entire simulation time
and at the four gauges. The Manning's roughness coefficient
in the bottom of ocean $N_{3}$ at gauge number 21419 exhibits small sensitivity index 
during the second hour of simulation and Manning's roughness coefficient
at the land $N_1$ appears to be an insignificant contributor
to the variance.

\comment{In 2D, the sensitivity analysis shows that .... Figure~\ref{fig:sens2d}(top row) shows
the PC mean water surface elevation for the considered computational
domain at three different times as indicated in the title of each panel.
The standard deviation is also shown in Figure~\ref{fig:mean2d}(bottom row)}.

\subsubsection{Response surface}
In addition to the static and analysis, the PC surrogate 
can be used to construct a response surface for the uncertain input parameters.
This is achieved by sampling the PC surrogate for different values of the germ $\xxi$ within the prior
range.
\comment{The impact of $N_2$ and $N_3$ for fixed $N_1=0.1025$ is illustrated in Figure~\ref{fig:response2}
and Figure~\ref{fig:response3}
for the area-averaged SST within the $42~km\times42~km$ analysis region; the
different panels illustrate the time evolution of the SST response
surfaces. 
Their most striking features are the relatively flat
horizontal contours on Sep~17 and Sep~18 indicating that SST depends
only mildly on $V_{\max}$ even during peak winds;
unsurprisingly these contours turn completely flat on Sep
19 (and afterwards) when the winds dip below 10~m/s. The multiplicative
factor $\alpha$, however, exerts a strong influence on SST even
under mild wind conditions.  This influence is relatively
weak for the low $\alpha$ range and increases substantially for the
higher $\alpha$ as evidenced by the packed contours.  The temperature
response surfaces at 50-m and 200-m depths (not shown) exhibit
roughly the same structure as the surface, with milder
dependence on $V_{\max}$ even during the peak winds of Sep~18.}

### 2013-11-12 paper/formulation.tex

\section{Formulation}
\label{sec:formu}
We begin by introducing the essential building blocks of the present formulation: 
the Bayesian approach to inverse problems~\ref{sec:inference}
and the stochastic spectral methods for forward uncertainty propagation in ~\ref{sec:uqpce}
\input{inference}
\input{uqpce}

>>>>

\section{Formulation}
\label{sec:formu}
We begin by introducing the methods employed in this work.
The Bayesian approach to inverse problems is described in Section~\ref{sec:inference};
and its application to our problem of inferring the Manning Friction Coefficients is described in Section~\ref{sec:manning}. Finally in Section~\ref{sec:uqpce}, we introduce
a stochastic spectral method for forward uncertainty propagation and accelerating the Bayesian inference.
%The two pieces are brought together in~\ref{sec:setup}.

\input{inference}
\input{coefs}

\input{uqpce}

### 2013-11-12 paper/pcstudy.tex
adding inference section
Fix figure paths

Figure~\ref{fig:rlzs} shows the evolution of the
water surface elevation for the 125 realizations obtained 
to compute the PC expansions at the four different gauge 
locations. We notice that the variation in surface elevation 
is negligible during the first hour at gauge number 21418 
and during the first two hours for the remaining gauges
as the plots of the different realizations are seen to superimpose.
However, a noticeable uncertainty in the surface 
elevation can be seen  at all gauges as indicated from the thickness 
of the bands formed by the plots of the realizations. This uncertainty
occurs till the end of the simulations.
\begin{figure}[h]
\centering
\begin{tabular}{clc}        
\includegraphics[width=0.6\textwidth]{figures/rlzs_gauges.pdf} 
\end{tabular}
\caption{125 Geoclaw realizations at different gauge locations.}
\label{fig:rlzs}
\end{figure}

In order to check the consistency of the approximation, we compare
 water surface elevation from the realizations 
with those obtained from the PC surrogate. The different curves
show an excellent agreement (not shown) for all times. We also define
an error metric that measures the relative normalized root mean-square error between the left hand side function 
in Equation~(\ref{eq:stochseries})
and its PC representation at the sampling points:
\begin{equation} 
   E = \frac{\displaystyle
         \left(\sum_{\xxi \in \NISP} \left|U(\xxi) - \sum_{k = 0}^{P}
U_k\Psi_k(\xxi)\right|^2
         \right)^{1/2}}
        {\displaystyle
          \left(\sum_{\xxi \in \NISP} \left|U(\xxi)\right|^2\right)^{1/2} 
          },
\label{eq:error}
\end{equation}
where $\NISP$ is the 125-member ensemble obtained using the PC algorithm. 
This error metric calculated at the different gauge locations is shown in Figure~\ref{fig:error};
the largest relative normalized error for 
water surface elevation is about 0.1\% . 
\begin{figure}[h]
\centering
\begin{tabular}{clc}        
\includegraphics[width=0.5\textwidth]{figures/error_gauge1.pdf} &
\includegraphics[width=0.5\textwidth]{figures/error_gauge2.pdf} \\
\includegraphics[width=0.5\textwidth]{figures/error_gauge3.pdf} &
\includegraphics[width=0.5\textwidth]{figures/error_gauge4.pdf}

\end{tabular}
\caption{125 Geoclaw realizations at different gauge locations.}
\label{fig:error}
\end{figure}

@@

\begin{figure}[h]
\centering

\begin{tabular}{clcl}
\includegraphics[width=0.5\textwidth]{figures/pdfs1_2.pdf} &
\includegraphics[width=0.5\textwidth]{figures/pdfs2_2.pdf} \\
\includegraphics[width=0.5\textwidth]{figures/pdfs3_2.pdf} &
\includegraphics[width=0.5\textwidth]{figures/pdfs4_2.pdf}
\end{tabular}
\caption{pdf of water surface elevation at the different gauge locations at t = 7200 s.}
\label{fig:pdfs2}
\end{figure}

\begin{figure}[h]
\centering
\begin{tabular}{clc}

\includegraphics[width=0.5\textwidth]{figures/pdfs1_3.pdf} &
\includegraphics[width=0.5\textwidth]{figures/pdfs2_3.pdf} \\
\includegraphics[width=0.5\textwidth]{figures/pdfs3_3.pdf} &
\includegraphics[width=0.5\textwidth]{figures/pdfs4_3.pdf}
\end{tabular}
\caption{pdf of water surface elevation at the different gauge locations at t = 10800 s.}
\label{fig:pdfs3}
\end{figure}
The different curves
correspond to increased order of PC, from 1 to 5; 
The plots indicate that the double peaked distributions are
well-resolved with order  4 but becomes weakly insenstive to further refinement 
as shown with order 5.

>>>>

Figure~\ref{fig:rlzs} shows the evolution of the
water surface elevation for the 125 realizations obtained 
to compute the PC expansions at the four different gauge 
locations. We notice that the variation in surface elevation 
is negligible during the first hour at gauge number 21418 
and during the first two hours for the remaining gauges
as the plots of the different realizations are seen to superimpose.
However, a noticeable uncertainty in the surface 
elevation can be seen  at all gauges as indicated from the thickness 
of the bands formed by the plots of the realizations. This uncertainty
occurs till the end of the simulations.

In order to check the consistency of the approximation, we compare
 water surface elevation from the realizations 
with those obtained from the PC surrogate. The different curves
show an excellent agreement (not shown) for all times. We also define
an error metric that measures the relative normalized root mean-square error between the left hand side function 
in Equation~(\ref{eq:stochseries})
and its PC representation at the sampling points:
\begin{equation} 
   E = \frac{\displaystyle
         \left(\sum_{\xxi \in \NISP} \left|U(\xxi) - \sum_{k = 0}^{P}
U_k\Psi_k(\xxi)\right|^2
         \right)^{1/2}}
        {\displaystyle
          \left(\sum_{\xxi \in \NISP} \left|U(\xxi)\right|^2\right)^{1/2} 
          },
\label{eq:error}
\end{equation}
where $\NISP$ is the 125-member ensemble obtained using the PC algorithm. 
This error metric calculated at the different gauge locations is shown in Figure~\ref{fig:error};
the largest relative normalized error for 
water surface elevation is about 0.1\% .

@@

The different curves
correspond to increased order of PC, from 1 to 5; 
The plots indicate that the double peaked distributions are
well-resolved with order  4 but becomes weakly insenstive to further refinement 
as shown with order 5.

### 2013-11-12 paper/tsunami.tex

Physical and computational parameters
Figure~\ref{fig:setup}(left) shows the topography of the considered domain.
Land topography is shown using while the bathymetry is shown using a lighter colormap.

\begin{figure}[h]
\centering
\begin{tabular}{clc}
\includegraphics[width=0.45\textwidth]{./figures/topo.pdf}  &
\includegraphics[width=0.45\textwidth]{./figures/coef.pdf} 
\label{setup}
\end{tabular}
\caption{Topography and gauge locations.}
\label{fig:setup}
\end{figure}

\subsection{Parametric Uncertainty}

\subsection{Observations}

Observations used in th

During this tsunami, different gauges were installed to observed the surface elevation
The locations of theses gauges are shown in Figure~\ref{setup} .
The gauges are used in the inverse problem to infer the manning coefficients.

\begin{figure}[h]      
\centering
\includegraphics[width=0.6\textwidth]{./figures/obs.pdf}
\caption{Observed data of water surface elevation with time at four different gauges.}
\label{fig:obs}
\end{figure}

>>>>

\subsection{Parametric Uncertainty}
Uniform distribution is assumed as the prior for the three manning coefficients.
such that UN = [0.005-0.2]. The three regions are shown in Figure~\ref{fig:setup}(right)

\subsection{Observations}
During this tsunami, different gauges were installed to observed the surface elevation
The locations of theses gauges are shown in Figure~\ref{setup} .
The gauges are used in the inverse problem to infer the manning coefficients.
